How to perform Hierarchical Clustering using `hclust`
========================================================

`hierarchical clustering` is a method of cluster analysis which seeks to build a hierarchy of clusters.  Strategies for hierarchical clustering generally fall into two types:

- **Agglomerative**: This is a "bottom up" approach: each observation starts in its own cluster, and pairs of clusters are merged as one moves up the hierarchy.
- **Divisive**: This is a "top down" approach: all observations start in one cluster, and splits are performed recursively as one moves down the hierarchy.



Draw a sample of 40 records from iris data, and remove variable Species

```{r}
idx  <- sample(1:dim(iris)[1],40)
irisSample  <- iris[idx,]
irisSample$Species  <- NULL
```

Hierarchial Clustering 
```{r fig.width=7, fig.height=6}
hc  <- hclust(dist(irisSample), method="ave")
plot(hc, hang = -1, labels = iris$Species[idx])
```


Reference 
- https://en.wikipedia.org/wiki/Hierarchical_clustering
