Human Activity Recognition Using Smartphones
========================================================
http://rpubs.com/wangwf/19053

One of the most exciting areas in all of data science right now is wearable computing - see for example this article.

Data set information was download from [http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones](http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones). 

"The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING UPSTAIRS, WALKING DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data. The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain.”

For each record in the raw dataset it is provided:


```
* Triaxial acceleration from the accelerometer (total acceleration) and the estimated body acceleration.
* Triaxial Angular velocity from the gyroscope.
* A 561-feature vector with time and frequency domain variables.
* Its activity label.
* An identifier of the subject who carried out the experiment.
```

Step#1 download and unzip files


```{r}
downloadFiles <- function(dataURL = "", destF = "t.csv") {
    if (!file.exists(destF)) {
        download.file(dataURL, destF, method = "wget")
    } else {
        message("data already downloaded.")
    }
}
```

```{r}
dataURL = "http://archive.ics.uci.edu/ml/machine-learning-databases/00240/UCI%20HAR%20Dataset.zip"
destF = "UCI HAR Dataset.zip"
downloadFiles(dataURL,destF)

```

```{r}
if (!file.exists("UCI HAR Dataset")) unzip(destF)
```

```{r}
system("ls")
```


```{r}
nsampleSize = -1L
train <- readData("UCI_HAR_Dataset/train", "train", nsampleSize)
test <- readData("UCI_HAR_Dataset/test", "test", nsampleSize)
```

```{r}
train <- transform(train, subjectID = factor(subjectID), activityID = factor(activityID))
test <- transform(test, subjectID = factor(subjectID), activityID = factor(activityID))

train$partition = "train"
test$partition = "test"
samsungData <- rbind(train, test)
summary(samsungData$subjectID)
```


This function loads a file <f> from a zip archive <zipfile>
Prints a message with the indented filename <f>
```{r}
load.with.msg <- function (zipfile, f) {
    message(sprintf("    \"%s\"", f))
    df <- read.table(unz(zipfile, f))
    df
}
```

```{r}
# Local file to store the compressed data
raw.data.file <- "UCI_HAR_Dataset.zip"

```



############################################################
# Loading data from zip file                               # 
############################################################
```{r}
message("\nLoading metadata...")
activity_labels <- load.with.msg(raw.data.file, zfilename.activity_labels)
features <- load.with.msg(raw.data.file, zfilename.features)

# Load training data
message("Loading training data (can take some time...)")
data.train <- lapply(load.list.train, 
                     function(f) load.with.msg(raw.data.file, f))


message("Loading test data (can take some time...)")
# Load test data
data.test <- lapply(load.list.test, 
                     function(f) load.with.msg(raw.data.file, f))

# Check if loaded trainign and test dataframes have same number of cols 
assertion.datacols <- identical(lapply(data.train, FUN=ncol), 
                                lapply(data.test, FUN=ncol))

message(sprintf("Checking if training and test data sets %s: %s", 
                "have same number of columns", 
                assertion.datacols))

if (!assertion.datacols) {
    stop("Training and test data sets must have same number of columns!")
}
```









```{r}
assertion.datanames <- identical(sort(names.train), sort(names.test))

message(sprintf("Checking if same files exist for training and test data: %s",
                assertion.datanames))

if (!assertion.datanames) {
    stop("Training and test data must have same file structure!")
}

```







