Linear Regression
========================================================


Create and Load data  
```{r}
year  <- rep(2008:2010, each=4)
quarter  <- rep(1:4, 3)
cpi  <- c(162.2, 164.6, 166.5, 166.0,
          166.2, 167.0, 168.6, 169.5,
          171.0, 172.1, 173.3, 174.0)
plot(cpi, xaxt="n", ylab="CPI", xlab="")
# draw x-axis
axis(1, labels=paste(year,quarter,sep="Q"), at=1:12, las=3)
```

Check correlation between CPI and other variable

```{r}
cor(year, cpi)
```


```{r}
cor(quarter, cpi)
```

build linear regression model with `lm()` function on this data, using year and quarter as predicators and CPI as response.

```{r}
fit  <- lm(cpi ~ year + quarter)
fit
```

Predict CPI in 2011

```{r}
(cpi2011 <- fit$coefficients[[1]] + fit$coefficients[[2]]*2011 +
   fit$coefficients[[3]]*(1:4))
```

more details about the model .

```{r}
attributes(fit)
```

```{r}
fit$coefficients
```

The differences between observed values and fitted values can be obtained with function `residuals()`.

```{r}
residuals(fit)
```

```{r}
summary(fit)
```

plot the model 


```{r}
par(mfrow=c(2,2))
plot(fit)
```


pltot the model in 3D
-------


- `scatterplot3d()` creates a 3D scatter plot and `plane3d()` draws the fitted plane.
- Parameter `lab` specifies the number of tickmarks on the x- and y-axes.

```{r}
par(mfrow=c(1,1))
library(scatterplot3d)
s3d  <- scatterplot3d(year, quarter, cpi, highlight.3d=T, 
                      type="h", lab=c(2,3))

s3d$plane3d(fit)
```


Predict cpi in 2011

```{r}
data2011  <- data.frame(year=2011, quarter = 1:4)
cpi2011  <- predict(fit, newdata = data2011)
style  <- c(rep(1,12), rep(2,4))
plot(c(cpi, cpi2011), xaxt="n", ylab="CPI", xlab="", pch=style, col=style)

axis(1, at=1:16, las=3,
     labels=c(paste(year,quarter,sep="Q"), "2011Q1", "2011Q2", "2011Q3", "2011Q4"))

```

